---
title: "Tight Ends"
output: html_document
---
<style type="text/css">

body {
  font-size: 14pt;
}

h1 { /* Header 1 */
  font-size: 34px;
  color: lightcoral;
  font-weight: bold;
}
</style>

[FantasyPros Draft Pick](https://www.fantasypros.com/nfl/rankings/te-cheatsheets.php){target="_blank"}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
```

```{r, include=FALSE}
library(here)
library(tidyverse) 
library(janitor)
library(httr)
library(jsonlite)
library(RCurl)
library(glue)
library(mysportsfeedsR)
library(kableExtra)
library(gt)
library(ggeasy)
library(shiny)
library(reactable)
library(ggrepel)

setwd(here("../"))
source("login_creds.R")
```

<!-- # API request -->
```{r, include=FALSE}
res = GET('https://api.mysportsfeeds.com/v2.0/pull/nfl/2020-2021-regular/player_stats_totals.json', 
          add_headers(Authorization = paste("Basic", auth)))
api_response <- content(res, as="text")
api_response <- jsonlite::fromJSON(api_response, flatten=TRUE)

player_stats <- api_response$playerStatsTotals %>%
  as_tibble() %>%
  clean_names()  %>%
  select(first_name = player_first_name, last_name = player_last_name, position = player_primary_position, current_team = player_current_team_abbreviation, games_played = stats_games_played, rush_yards = stats_rushing_rush_yards, rush_td = stats_rushing_rush_td, receive_targ = stats_receiving_targets, receive_rec = stats_receiving_receptions, receive_yards = stats_receiving_rec_yards, receive_avg = stats_receiving_rec_average, receive_td = stats_receiving_rec_td, receive_lng = stats_receiving_rec_lng, receive_20plus = stats_receiving_rec20plus, receive_40plus = stats_receiving_rec40plus)
```


- Indexed games so everyone is scored based on playing all 16 games
- `Golden Ratio`: a composite score including:
   - Receiving yards 42% weight 
   - Receiving td 42% weight 
   - Total receptions 16% weight
   - rush yards not included because max was only 26 yards
- Top 50
- Rookies: 
  - Kyle Pitts (ATL) \#1 draft pick 2021 (Tier 1: Rank 5)
```{r}
wr_stats_explore <- player_stats %>%
  select(first_name, last_name, position, current_team, games_played, receive_yards, receive_td, receive_rec, rush_yards, rush_td) %>%
  filter(position == "TE") %>%
   select(-position)

# range(wr_stats_explore$rush_td)
# range(wr_stats_explore$rush_yards)

te_stats <- player_stats %>%
  select(first_name, last_name, position, current_team, games_played, receive_yards, receive_td, receive_rec) %>%
  filter(position == "TE") %>%
   select(-position) %>%
  mutate(games_index = games_played/16) %>%
  relocate(games_index, .before = receive_yards) %>%
  mutate(full_name = paste(first_name, last_name),
         receive_yards = receive_yards/games_index, 
         receive_td = receive_td/games_index, 
         receive_rec = receive_rec/games_index,
         receive_yards_i = receive_yards/max(receive_yards),
         receive_td_i = receive_td/max(receive_td), 
         receive_rec_i = receive_rec/max(receive_rec),
         golden_ratio = (receive_yards_i*1.42)+(receive_td_i*1.42)+(receive_rec_i*1.16)) %>%
  arrange(desc(golden_ratio)) %>%
  select(full_name, last_name, current_team, games_played, receive_yards, receive_td, receive_rec, golden_ratio, full_name) %>%
  mutate(rank = 1:length(golden_ratio)) %>%
  relocate(rank) %>%
  filter(rank <= 50) %>%
  mutate_if(is.numeric, round, 2) 



te_stats_picks <- te_stats %>%
  select(-last_name) %>%
  rename(Rank = rank, 
         "Name" = full_name,
         Team = current_team,
         "Games Played" = games_played,
         "Receiving Yards" = receive_yards, 
         "Receiving TD" = receive_td,
         "Receive Receptions" = receive_rec,
         "Golden Ratio" = golden_ratio)

#write.csv(te_stats_picks, "nfl/wr_picks.csv", row.names=FALSE)

reactable(te_stats_picks, selection = "multiple", defaultPageSize = 50, highlight=TRUE, striped=TRUE)

```

# Receiving TDs and Yards
```{r}
te_stats_plot <- te_stats %>%
  filter(rank < 31)

ggplot(te_stats_plot, aes(receive_td, receive_yards, label = last_name)) +
  geom_point() +
  geom_text_repel() +
  theme_bw() +
  labs(x="Receiving TDs", y="Receiving Yards") +
  easy_text_size(15)

```

<br>
```{r}
ggplot(te_stats, aes(receive_yards)) +
  geom_histogram(bins = 20, fill="darkblue") +
  theme_bw() +
  labs(x="Receiving Yards") +
  easy_text_size(15)
```

<br>
```{r}
ggplot(te_stats, aes(receive_td)) +
  geom_histogram(bins = 20, fill="darkblue") +
  theme_bw() +
  labs(x="Receiving TDs") +
  easy_text_size(15)
```


# \# of Receptions
```{r}
ggplot(te_stats, aes(receive_rec)) +
  geom_histogram(bins = 20, fill="darkblue") +
  theme_bw() +
  labs(x="Number of Receptions") +
  easy_text_size(15)
```




